---
title: "Practical Machine Learning Project"
author: "Toh KC"
date: "November 4, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Data Preparation
Loading required libraries and preparing data sets
```{r prep}
library(caret)
library(randomForest)
training <- read.csv("pml-training.csv", na.strings = c("NA", ""))
testing <- read.csv("pml-testing.csv", na.strings = c("NA", ""))
trainingcheck<-dim(training)
testingcheck<-dim(testing)
```
Training data consist of 19622 observations and 160 variables
Testing data consist of 20 observations and 160 variables

## Data Cleaning
Removing data null values and selecting features
```{r remove}
training2 <- training[, colSums(is.na(training)) == 0]
testing2 <- testing[, colSums(is.na(testing)) == 0]
training2<- training2[, -c(1:7)]
testing2 <- testing2[, -c(1:7)]
trainingcheck2<-dim(training2)
```
The cleaned data sets have 53 variables. Number of observations remain unchanged

## Data Grouping
Splitting the data into 2 groups, for training and validation
```{r group}
set.seed(1234) 
inTrain <- createDataPartition(training2$classe, p = 0.7, list = FALSE)
train <- training2[inTrain, ]
validation <- training2[-inTrain, ]
```

## Prediction
Selection of random forest for the modelling
```{r randomforest}
fit <- randomForest(classe ~ ., data = train, importance = TRUE, ntrees = 10)
model <- predict(fit, train)
print(confusionMatrix(model, train$classe))
accuracy <- postResample(model, train$classe)
accuracy

```
Accuracy is 100%. The model is suitable. Next step will be to test the model with the validation data

## Validation
```{r validation}
modelv <- predict(fit, validation)
print(confusionMatrix(modelv, validation$classe))
accuracyv <- postResample(modelv, validation$classe)
accuracyv

```
Accuracy is 99.6%. This model can be used with the test data set

## Test
Testing the model with the test data
```{r test}
modeltest <- predict(fit, testing2)
modeltest

```